---
title: "Darwin Core mapping"
subtitle: "For: Monitoring of fishes and crustaceans by the Province East Flanders, Belgium"
author:
- Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load packages:

```{r message = FALSE}
library(readr)          # To read data
library(dplyr)          # To do data wrangling
library(tidylog)        # To provide feedback on dplyr functions
library(here)           # To find files
library(sf)             # To convert coordinate systems
```

# Read source data

Create a data frame `occurrences` from the source data:

```{r read_inputs}
occurrences <- readr::read_csv(
  file = here::here("data", "raw", "vissen_en_crustacea.csv"),
  col_types = cols(.default = col_character())
) 
```

Preview data:

```{r preview_input_data}
occurrences %>% head(n = 5)
```

## Change coordinate system

Data are provided in Belgian Lambert72 ([EPSG 31370](https://epsg.io/31370)) coordinates. DwC accepts data in WGS84 ([EPSG 4326](https://epsg.io/4326)) only.

```{r add_wgs84_coords}
occurrences <- 
  occurrences %>% 
  st_as_sf(crs = st_crs(31370), wkt = "Shape") %>%
  st_transform(crs = 4326)
coords <- as_tibble(st_coordinates(occurrences))
occurrences <- 
  as_tibble(occurrences) %>% 
  bind_cols(coords) %>%
  select(-Shape)
```

Preview:

```{r}
occurrences %>% head(5)
```

# Create database

Create a SQLite database with the source data, so it can be queried with SQL in the next steps:

```{r create_db}
message("Create in memory SQLite database...")
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
# import occurrences
DBI::dbWriteTable(con, "occurrences", occurrences)
message("DONE")
```

# Darwin Core mapping

Create [Event](https://rs.gbif.org/core/dwc_event_2022-02-02.xml) core:

```{r event}
dwc_event_sql <- glue::glue_sql(
  readr::read_file(here::here("sql", "dwc_event.sql")),
  .con = con
)
dwc_event <- 
  DBI::dbGetQuery(con, dwc_event_sql) %>%
  dplyr::arrange(as.double(gsub("VIS_", "", eventID)))
```

Create [Occurrence](https://rs.gbif.org/core/dwc_occurrence_2022-02-02.xml) extension:

```{r occurrence}
message("Map occurrences to DwC...")
dwc_occurrence_sql <- glue::glue_sql(
  readr::read_file(here::here("sql", "dwc_occurrence.sql")), 
  .con = con
)
dwc_occurrence <- 
  DBI::dbGetQuery(con, dwc_occurrence_sql) %>% 
  dplyr::arrange(as.double(gsub("VIS_", "", occurrenceID)))
message("DONE")
```

Create [Measurement Or Facts](https://rs.gbif.org/extension/dwc/measurements_or_facts_2022-02-02.xml) extension:

```{r mof}
dwc_mof_sql <- glue::glue_sql(
  readr::read_file(here::here("sql", "dwc_mof.sql")), 
  .con = con
)
dwc_mof <- DBI::dbGetQuery(con, dwc_mof_sql)%>% 
  dplyr::arrange(as.double(gsub("VIS_", "", eventID)))
```

# Save data to CSV

```{r save_csv}
write_csv(dwc_event, here::here("data", "processed", "event.csv"), na = "")
write_csv(
  dwc_occurrence, here::here("data", "processed", "occurrence.csv"),
  na = ""
)
write_csv(dwc_mof, here::here("data", "processed", "mof.csv"), na = "")
```

# Overview changes

In this section we summarize the changes in the DwC output after the very last udpate.

## Read DwC files

Read DwC mapping files from both `main`:

```{r main}
dwc_event_main <- readr::read_csv(
  file = "https://raw.githubusercontent.com/riparias/pov-fishes-occurrences/main/data/processed/event.csv",
  col_types = cols(.default = "c"))
dwc_occurrence_main <- readr::read_csv(
  file = "https://raw.githubusercontent.com/riparias/pov-fishes-occurrences/main/data/processed/occurrence.csv",
  col_types = cols(.default = "c"))
dwc_mof_main <- readr::read_csv(
  file = "https://raw.githubusercontent.com/riparias/pov-fishes-occurrences/main/data/processed/mof.csv",
  col_types = cols(.default = "c"))
```

and actual branch (`automatic-update-*`):
 
```{r automatic-update}
dwc_event_update <- readr::read_csv(
  file = here::here("data","processed","event.csv"),
  col_types = cols(.default = "c"))
dwc_occurrence_update <- readr::read_csv(
  file = here::here("data","processed","occurrence.csv"),
  col_types = cols(.default = "c"))
dwc_mof_update <- readr::read_csv(
  file = here::here("data","processed","mof.csv"),
  col_types = cols(.default = "c"))
```

## Overview

How many new events (`eventID`) have been added in event core?

```{r new_eventID_values}
new_events <- 
  dwc_event_update %>%
  filter(!eventID %in% dwc_event_main$eventID)
message("New events:")
print(new_events)
```

How many occurrences (`occurrenceID`) have been added in occurrence extension?

```{r new_occurrenceID_values}
new_occs <- 
  dwc_occurrence_update %>%
  filter(!occurrenceID %in% dwc_occurrence_main$occurrenceID)
message("New occurrences:")
print(new_occs)
```

Have some events been removed?

```{r removed_events}
removed_events <- 
  dwc_event_main %>%
  filter(!eventID %in% dwc_event_update$eventID)
message("Removed events:")
print(removed_events)
```


Have some occurrences been removed?

```{r removed_occs}
removed_occs <- 
  dwc_occurrence_main %>%
  filter(!occurrenceID %in% dwc_occurrence_update$occurrenceID)
message("Removed occurrences:")
print(removed_occs)
```

Total number of rows in the new event core in comparison with the previous version:

```{r nrows_events}
message(paste("Number of events (new):", nrow(dwc_event_update)))
message(paste("Number of events (old):", nrow(dwc_event_main)))
message(paste("Difference:", 
              nrow(dwc_event_update) - nrow(dwc_event_main))
)
```

Total number of rows in the new occurrence extension in comparison with the previous version:

```{r nrows_occs}
message(paste("Number of occurrences (new):", nrow(dwc_occurrence_update)))
message(paste("Number of occurrences (old):", nrow(dwc_occurrence_main)))
message(paste("Difference:", 
              nrow(dwc_occurrence_update) - nrow(dwc_occurrence_main))
)
```

Total number of rows in the new measurement or facts (mof) extension in comparison with the previous version:

```{r nrows_mof}
message(paste("Number of mofs (new):", nrow(dwc_mof_update)))
message(paste("Number of mofs (old):", nrow(dwc_mof_main)))
message(paste("Difference:", 
              nrow(dwc_mof_update) - nrow(dwc_mof_main))
)
```

## New species

### New `scientificName`
The DwC `scientificName` is a manual mapping of scientific names from field `ScientificName`. It is therefore important to get a list with all new values to be added to the mapping:

```{r new_Sporen_Waarnemingen_Naam}
occs_without_sc_name <- 
  dwc_occurrence_update %>%
  filter(is.na(scientificName)) %>%
  pull(occurrenceID)
query <- glue::glue_sql("
  SELECT 
    DISTINCT o.\"ScientificName\" AS \"ScientificName\"
  FROM occurrences AS o 
  WHERE  
  o.\"ObservationIdentifier\" IN ({occs_without_sc_name*})
  ", .con = con)
query
raw_sc_names <- dplyr::as_tibble(DBI::dbGetQuery(con, query))
message("New values in field 'ScientificName' to be mapped:")
print(raw_sc_names)
```

### New `vernacularName`

The DwC `vernacularName` is a manual mapping of Dutch vernacular names from field `DutchName`. It is therefore important to get a list with all new values to be added to the mapping:

```{r new_DutchName}
occs_without_vernacular_name <- 
  dwc_occurrence_update %>%
  filter(is.na(vernacularName)) %>%
  pull(occurrenceID)
query <- glue::glue_sql("
  SELECT 
    DISTINCT o.\"DutchName\" AS \"Dutchname\"
  FROM occurrences AS o 
  WHERE  
  o.\"ObservationIdentifier\" IN ({occs_without_vernacular_name*})
  ", .con = con)
query
raw_vernacular_names <- dplyr::as_tibble(DBI::dbGetQuery(con, query))
message("New values in field 'DutchName' to be mapped:")
print(raw_vernacular_names)
```

Disconnect from SQLite database:

```{r disconnect}
DBI::dbDisconnect(con)
```
