---
title: "Fetch invasive alien plants data from POV"
author:
- Damiano Oldoni
- Sanne Govaert
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

Load packages:

```{r}
library(httr) # generic webservice package
library(ows4R) # interface for OGC webservices
library(here) # to work with paths
```

# Fetch data from WFS

WFS link:

```{r link}
wfs <- "https://gis.oost-vlaanderen.be/server/services/GEODATA/Geodata_Oost_Vlaanderen/MapServer/WFSServer"
```

Create WFS client:

```{r create_client}
pov_client <- ows4R::WFSClient$new(wfs, serviceVersion = "2.0.0")
```

Overview of the layers available:

```{r layers}
layers <- pov_client$getFeatureTypes(pretty = TRUE)
print(layers)
```

Request data from layer `Geodata_Oost_Vlaanderen:Waarnemingen_vissen_en_crustacea`:

```{r request}
url <- httr::parse_url(wfs)
url$query <- list(service = "wfs",
                  request = "GetFeature",
                  typename = "Geodata_Oost_Vlaanderen:Waarnemingen_vissen_en_crustacea",
                  outputFormat = "CSV"
)
request <- httr::build_url(url)
```

Download requested data as csv file and save them in `vissen_en_crustacea.csv`:

```{r download_data}
file_path <- here::here("data", "raw", "vissen_en_crustacea.csv")
httr::GET(url = request, write_disk(file_path, overwrite = TRUE))
```

## Remove column `GmlID`

The column `GMlID` is not present in the database of POV: it is added by the WFS service as a progressive number with the layer name as prefix. Removing this column from the raw data would increase the effectiveness of the git file change history tool, e.g. in case new observations are added at a certain row in the file.

```{r remove_GmlID}
# read raw occurrences
raw_occurrences <- readr::read_csv(
  file = file_path,
  na = "",
  col_types = readr::cols(.default = readr::col_character())
)

# remove GmlId column
raw_occurrences <- subset(raw_occurrences, select = -GmlId)

# save raw occurrences (overwrite file)
readr::write_csv(
  x = raw_occurrences,
  file = file_path,
  na = "",
)
```
